---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

### Set environment and load in data

```{r}
# Please add your main working directory
main_wd = ''
setwd(main_wd)
rm(list = ls())
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)
library(rpart)
library(class)
library(ROCR)
library(rpart.plot)
library(reshape2)
library(nnet)

df_order =read.csv("data/order.csv", header = TRUE, sep = ",")
df_online = read.csv("data/online.csv", header = TRUE, sep = ",")


```

#### *Analyze shape of Order and Online Dataframe, then print first couple rows*
###**1) Order DataFrame:**  

```{r}
#Rows x Columns
dim(df_order)
```  
`#Print Order Dataframe`
```{r echo=FALSE}
head(df_order)
```  
###**2) Online DataFrame:**  
```{r }
#Rows x Columns
dim(df_online)
```
`#Print Order Dataframe`
```{r echo=FALSE}
head(df_online)
```  
  

# **Exploratory Analysis**
## ***Order DataFrame***

From the Order dataframe we can see there are 6 variables.  
```{r echo=FALSE}
summary(df_order)

#df_order$custno <- as.factor(df_order$custno)
#df_order$ordno <- as.factor(df_order$ordno)
#df_order$prodcat2 <- as.factor(df_order$prodcat2)
#df_order$prodcat1 <- as.factor(df_order$prodcat1)

```

###--Column: Custno
**Customer_no56** is observed to have the highest count at **626**  
**Customer_no1** is observed to have the lowest count at **1**  
`Frequency Table`
```{r echo=FALSE}
head(summary(as.factor(df_order$custno)))
custnoCount <- table(df_order$custno)
head(custnoCount[order(custnoCount, decreasing = FALSE)])
rm(custnoCount)
```
```{r}
# Count of Unique Customers
length(unique(as.factor(df_order$custno)))
```




###--Column: Ordno

```{r}
# Total Unique Order Numbers
length(unique(df_order$ordno))
```

**order_no6076** has the highest count of obsrvations at **26** counts  
**order_no1** has lowest count of observations at **1**  
`Frequency Table`
```{r echo= FALSE}
head(summary(as.factor(df_order$ordno)))
orderCount <- table(df_order$ordno)
head(orderCount[order(orderCount, decreasing = FALSE)])
```




###--Column: Orderdate
First date time observed **2016-01-01 05:05:14**  
Last date time observed **2019-01-02 23:54:58 PST**
```{r echo=FALSE}
### change Orderdate from int to posixct format
df_order$orderdate <- as.POSIXct(df_order$orderdate)
summary(df_order)
```






###--Column: prodcat2

```{r }
# Total Category NA's
tail(summary(df_order$prodcat2))
```

```{r }
# Total Unique Values
length(unique(df_order$prodcat2))
```
**value_3** has the highest count of observations at **40,908**  
**value_194** has lowest count of observations at **1**  
```{r echo=FALSE}
head(summary(as.factor(df_order$prodcat2)))
proCat2Count <- table(df_order$prodcat2)
head(proCat2Count[order(proCat2Count, decreasing = FALSE)])
```



###--Column: prodcat1
Product Category includes the following 6 categories **1, 2, 3, 4, 5, 7**  
**Category 2** has highest count of orders at **88684**  
**Category 5** has lowest count of orders at **11180**  
```{r echo=FALSE}
summary(df_order$prodcat1)
length(unique(df_order$prodcat1))
ggplot(df_order, aes(prodcat1, fill = as.factor(prodcat1))) + geom_bar()

```




###--Column: revenue
Max revenue made from orders is **$150** and Min is **$0**  
**Average** made from order is **$74.94**  
**1st and 3rd Quartile range** are between **$37.43 and $112.28** 
```{r echo=FALSE}
summary(df_order$revenue)
```
 




##***Online DataFrame***
From the Online Dataframe we can observe 7 variables
```{r echo=FALSE}
summary(df_online)

#df_online$session <- as.factor(df_online$session)
#df_online$visitor <- as.factor(df_online$visitor)
#df_online$custno <- as.factor(df_online$custno)
#df_online$category <- as.factor(df_online$category)
#df_online$event1 <- as.factor(df_online$event1)
#df_online$event2 <- as.factor(df_online$event2)
```


###--Column: session
Highest count of sessions observed is **5** from **session_id72733**  
`Frequency Table`
```{r echo=FALSE}
head(summary(as.factor(df_online$session)))
```
Lowest count of sessions observed is **1**  
`Frequency Table`
```{r echo=FALSE}
seshCount <- table(df_online$session)
head(seshCount[order(seshCount, decreasing = FALSE)])
```

```{r}
# Total unique number of sessions
length(unique(df_online$session))
```



###--Column: visitor

```{r}
# Total unique number of visitors
length(unique(df_online$visitor))
```
***Lowest count*** of visitors are **1** with ***Max count*** at **440** visits from **visitor_ID328620**  
`Frequency Table`
```{r echo=FALSE}
head(summary(as.factor(df_online$visitor)))

vizCount <- table(df_online$visitor)
head(vizCount[order(vizCount, decreasing = FALSE)])

```
  




###--Column: dt
First date time observed **2016-01-01 00:00:08**  
Last date time observed **2017-12-31 23:58:05**
```{r echo=FALSE}
###change dt from int to posixct format
df_online$dt <- as.POSIXct(df_online$dt)
summary(df_online$dt)
```



###--Column: custno
 
```{r}
# Total unique number of customer number
length(unique(df_online$custno))
```
**cust_no713** is observed to have the highest counts at **855** 
**cust_no15** is observed to have lowest count at **1**  
`Frequency Table`
```{r echo=FALSE}
head(summary(df_online$custno))
custCount <- table(df_online$custno)
head(custCount[order(custCount, decreasing = FALSE)])
```



###--Column: category
**Category 3** has highest browsing count at **706920**  
Followed by **category 2** with **194890** counts  
**category 1** has least count of browsings with **52964**

```{r echo=FALSE}
table(df_online$category)
ggplot(df_online, aes(category, fill = as.factor(category))) + geom_bar()+
  scale_y_continuous( labels = comma)
```




###--Column: event1
**9** unique event_types for event1  
We observe **789846** *NA's* for Online Event 1  
For event1 **value 1** has the highest count of observations at **51567** counts  
For event1 **value 10** has the Lowest count of observations at **1223** counts  
```{r echo=FALSE}
summary(df_online$event1)
ggplot(df_online[!is.na(df_online$event1),], aes(event1, fill = as.factor(event1))) + geom_bar()+
  scale_y_continuous( labels = comma)
```
 



###--Column: event2
```{r}
# Total unique event2 types
length(unique(df_online$event2))
```

For event2 **value 7** has the highest counts at **367857** counts  
For event2 **value 10** has the lowest count at **8235** counts  

```{r echo=FALSE}
summary(df_online$event2)
table(df_online$event2)

ggplot(df_online, aes(event2, fill = as.factor(event2))) + geom_bar()+
  scale_y_continuous( labels = comma)
```
 




# **Feature Engineering and Feature Selection**



### Starting with fresh data set (after data exploration and design notes)
```{r}
rm(list = ls())
df_order =read.csv("data/order.csv", header = TRUE, sep = ",")
df_order <- df_order[df_order$prodcat1 %in% c(1,2,3),]
df_online = read.csv("data/online.csv", header = TRUE, sep = ",")



```

### Transforming data for feature selection, processing and merging
```{r}
# simplfying order by key
df_order$count <- 1
new_order <- aggregate(cbind(count) ~ custno + prodcat1, data = df_order, FUN = sum)

# simplfying online by key
df_online$count <- 1
df_online$event2 <- as.character(df_online$event2)

new_online <- aggregate(cbind(count) ~ custno + category + event2, data = df_online, FUN = sum)
new_online$event2 <- paste0("event_", new_online$event2)

# Feature transformation
new_online <- dcast(new_online, custno + category ~ event2, FUN = sum)
new_online[ is.na( new_online ) ] <- 0


names(new_order) <- c("custno", "category", "purchase_count")

# Merging data (only including order data)
main_df <- merge(new_order, new_online, by = c("custno", "category"), all.x = TRUE)

# QA
dim(main_df)
dim(new_order)
head(main_df)
```

### cleaning up column names and removing NA from data

```{r }

main_df <-main_df[, c("custno", "category", "purchase_count", paste0("event_", as.character(seq(10))))]
main_df[is.na(main_df)] <- 0


```




# **Model Design & Sampling**

###Create Test and Train Datasets 

```{r}
# setting up sample test and training
sample_d <- sample(seq(nrow(main_df)),  size= nrow(main_df)*.8, replace = FALSE)

train_df <- main_df[sample_d, c("category", paste0("event_", as.character(seq(10))))]
test_df <- main_df[-sample_d, c("category", paste0("event_", as.character(seq(10))))]
```

```{r }
# Test and prediction result
multi_log_reg <- multinom(category ~ ., data = train_df)
prediction_values <-predict(multi_log_reg, test_df[, c(paste0("event_", as.character(seq(10))))]) %>%
   as.character() %>% as.numeric()
summary(multi_log_reg)
```
# **Model Evaluation**


```{r}

#Counts Classified Incorrectly vs Correctly
performance <- table(prediction_values == test_df$category)
performance
#Accuracy Percentage
performance[2]/nrow(test_df) * 100
```

# **Summary of results**

From the Order Dataframe I normalized order in order to quanitfy the orders by customer and   category. From the Online Dataframe I normalized the event2 variable to featurize the variable Using these two new datasets I have transformed I then merge both datasets on 'custno' and 'category'. During my transformations and model building, I dropped the rows were product category was not 1, 2, or 3, from our order dataset. This was due there being no online.data for these product categories. This is how I reached my master dataset I would use to build my test and train samples.  
  
In order to predict the category a customer is to make a purchase from I created a 
dataset which contained the quanity of the category purchased, quantified the number of times event2 occured, category user browsed, and customerID number. A multinomial logistic regression is like a logistic regression but for classification tasks of two or more, which is ideal for my challenge.  




